m---
title: "BUA466_Project"
author: Team 8
date: "2023-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(PreProcess)
library(caret)
library(FNN)
library(lubridate)
library(fastDummies)
library(ggplot2)
library(randomForest)
library(rpart)
library(rpart.plot)
library(gains)
library(pacman)
library(plotly)
library(RColorBrewer)

```

```{r}
df <- read.csv("~/Library/CloudStorage/GoogleDrive-xwang7031@gmail.com/My Drive/BUA466:Project/data.csv")

```

# Familiar with data
## Missing value

```{r familar with data}

names(df)
str(df)

table(df$Car_Cancellation)
#cancel -743 
#not_cancel-9257 

#missing value
summary(df)
#package_id (8248/10000)

#to_area_id    (2091/10000)

#from_city_id   (6294/10000)
#to_city_id (9661/10000)

#from_data (0/10000)
#to_data.   (4178/10000)
#booking_created (0/10000)
```


# Data cleaning 
## Change to categorical variable

```{r}
#change online_booking, mobile_site_booking to dummy variable
#df <- dummy_cols(df,select_columns = c("online_booking","mobile_site_booking"))

#travel_type_id to Dummy variable
table(df$travel_type_id)
df$Longdistance <- ifelse(df$travel_type_id==1,1,0)
df$PointToPoint <- ifelse(df$travel_type_id==2,1,0)
df$HourlyRental <- ifelse(df$travel_type_id==3,1,0)

## Booking types____Traditional booking
df<-df%>%
  mutate(traditional_booking= ifelse(online_booking == 0 & mobile_site_booking == 0,1,0))

#from_area_id to categorical variable
df$from_area_id <-as.factor(df$from_area_id)

#to_area_id to categorical variable
df$to_area_id  <-as.factor(df$to_area_id )


#from_city_id  to categorical variable
df$from_city_id <-as.factor(df$from_city_id)
unique(df$from_city_id) #1, 15, NA

#to_city_id    to categorical variable
df$to_city_id <-as.factor(df$to_city_id)

```
## Waiting time

```{r waiting time}
#date(From_date, to_data, booking_created)
#########From_date
data_frame_F <- data.frame(datetime = mdy_hm(df$from_date))          
# extracting time
data_frame_F$from_time <- format(as.POSIXct(data_frame_F$datetime),format = "%H:%M")
# extracting date
data_frame_F$from_date <- as.Date (data_frame_F$datetime)


#########To_date
data_frame_T <- data.frame(datetime = mdy_hm(df$to_date))
# extracting time
data_frame_T$to_time <- format(as.POSIXct(data_frame_T$datetime),format = "%H:%M")
# extracting date
data_frame_T$to_date <- as.Date (data_frame_T$datetime)


#########booking_created
data_frame_B <- data.frame(datetime = mdy_hm(df$booking_created))
# extracting time
data_frame_B$created_date <- format(as.POSIXct(
  data_frame_B$datetime),format = "%H:%M")
# extracting date
data_frame_B$created_time <- as.Date (data_frame_B$datetime)

df_2<-df
df_2$from_date <-data_frame_F$datetime
df_2$to_date <-data_frame_T$datetime
df_2$booking_created<-data_frame_B$datetime

#sum(is.na(df_1$from_date)) =0
#sum(is.na(df_1$to_date)) =4178 $$$ we can delete to_data since it has so many missing value
#sum(is.na(df_1$booking_created)) =0


#from_date-booking_created=waiting time (make it positive)
df_1<-df_2%>%
  mutate(Waiting=difftime(from_date,booking_created,units = "hours"))

df_1$Waiting<-abs(df_1$Waiting)

##### create two new variable (from_time&from_date) for future analysis 
df_1<-df_1%>%
  mutate(from_time=data_frame_F$from_time)%>%
  mutate(from_date =data_frame_F$from_date )

```

```{r}
##### change from_data to (month, wkday, quart) for future analysis
df_1<- df_1|>
  mutate(month = month(from_date, label=T, abbr=T),
         wkday = wday(from_date, label=T, abbr=T),
         quart = quarter(from_date))

#change (month, wkday) to factor variables
df_1 <- df_1|> 
  mutate(month = factor(month,     
                         levels=c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                                  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")),
         wkday = factor(wkday,     
                         levels=c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"),
                         labels= c("M", "T", "W", "Th", "F", "Sa", "Su"))) 
```

# Descriptive analysis (3 plots)
## Barplot for Car_Cancellation & vehicle_model_ids

```{r}
# Barplot for Car_Cancellation & vehicle_model_ids
table <- as.data.frame.matrix(table(df$vehicle_model_id,df$Car_Cancellation))
table$model <-rownames(table)
table <- data.frame(table) ###

data <- data.frame(x = table$`1`, y1 = table$`0`)

table.df <- data.frame(
  'Model' = rownames(table),
  'Accepted' = table[,1],
  'Canceled' = table[,2])

Car_model_plot <- ggplot(data = table.df) +
  geom_bar(aes(x = Model, y = Accepted, fill = "Accepted"), stat = "identity") +
  geom_bar(aes(x = Model, y = Canceled, fill = "Canceled"), stat = "identity") +
  xlab("Model") + ylab("Number of Cars") +
  scale_fill_manual(values = c("Accepted" = "#7a86ec", "Canceled" = "orange"), name = "", guide = guide_legend(reverse=TRUE)) +
  ggtitle("Car Model Distribution")

interactive_car_plot <- ggplotly(Car_model_plot)
interactive_car_plot


```

## Barplot for Car_Cancellation & travel type 

```{r}
#Barplot for Car_Cancellation & travel type 

travel_type_id <- factor(df_1$travel_type_id)
counts <- table(travel_type_id)

table_travel_type<- table(df$travel_type_id , df$ Car_Cancellation)
barplot(table_travel_type, main="Travel Type and Car Cancellation",
  xlab="Travel Types", col=c("blue","#7a86ec","orange"),
  legend = rownames(counts), beside=TRUE)

```

## Plots for Car_Cancellation & time (months & weeks)

```{r}
# Bar plot for Car_Cancellation & time (months)
(month_cancellation  <- df_1 |> 
   ggplot() + 
  geom_bar(aes(x=month, y=Car_Cancellation, fill=wkday), 
           stat="identity")+
    theme_classic())

# Line plot for Car_Cancellation & time (weeks)
daily_cancellation <- df_1 %>%
  filter(Car_Cancellation == 1) %>%
  group_by(wkday) %>%
  summarise(cancel_count = n())

week_cancellation <- ggplot(data = daily_cancellation, aes(x = wkday, y = cancel_count)) +
  geom_line(color = "blue") +
  geom_point(size = 3, color = "black", shape = 20) +
  theme_bw() +
  labs(x = "Weekday", y = "Car Cancellation Count(=1)")

week_cancellation

```

#Predictive analysis & validation 

```{r}
# create a new data set for model
df_2<-df_1%>%
  select(#vehicle_model_id, user_id,
         Longdistance,
         PointToPoint,
         HourlyRental,
         month,
         wkday,
         Waiting,
         online_booking,
         mobile_site_booking,
         traditional_booking,
         Dist_trip_length,
         Car_Cancellation) 

#to factor
#df_2$user_id<-as.factor(df_2$user_id)
#df_2$vehicle_model_id<-as.factor(df_2$vehicle_model_id)

#to numeric
df_2$Waiting <- as.numeric(df_2$Waiting)
df_2$month<-as.numeric(df_2$month)
df_2$wkday<-as.numeric(df_2$month)

str(df_2)

```
# Oversampling
```{r}
#Oversampling the train data for Modal_2
set.seed(111)
train.s  <- sample(row.names(df_2),0.6*dim(df_2)[1] , prob = ifelse(df_2$Car_Cancellation>0, 0.9, 0.01)) 

train.s.df <- df_2[train.s, ] 

```

# Split data
```{r}
#Set 60% for training, 40% for validation
set.seed(111)
train <- sample(row.names(df_2), 0.6*dim(df_2)[1])  
valid <- setdiff(row.names(df_2), train)  

train.df <- df_2[train, ] 
valid.df <- df_2[valid, ] 

```


```{r}
# normalize data
train.df.norm<- preProcess(train.df, method="range")
valid.df.norm <- preProcess(valid.df , method="range")

train.df.norm<-predict(train.df.norm, train.df )
valid.df.norm <-predict(valid.df.norm , valid.df)

na.omit(train.df.norm)
na.omit(valid.df.norm)

```

# Model_1 - Logistic regression model

```{r}
#Modal_1 - Logistic regression model

logit.reg <- glm(Car_Cancellation~., data=train.df.norm, family = "binomial")

summary(logit.reg)
```

```{r}
# stepwise selection for variables
step(logit.reg, direction = "both")

# revised logistic regression model 
logit.reg2 <-  glm(formula = Car_Cancellation ~ Longdistance + PointToPoint + 
    month + online_booking + mobile_site_booking, family = "binomial", 
    data = train.df)

summary(logit.reg2)
#The Logit model can be written as:
#Car_Cancellation=-4.67131+
#( -0.86950*Longdistance)+
#(-0.86950*PointToPoint)+
#(0.16819*month)+
#(1.16568*online_booking)+
#(1.16123*mobile_site_booking)

```


## Model_1 - Model validation 
```{r}
# evaluate performance of logistic regression model
logit.reg.pred.train <- predict(logit.reg, train.df.norm, type = "response")
logit.reg.pred.valid <- predict(logit.reg, valid.df.norm, type = "response")


val.reg.train<-data.frame(actual = train.df.norm$Car_Cancellation, predicted = logit.reg.pred.train, errors =  logit.reg.pred.train - train.df.norm$Car_Cancellation)


val.reg.valid<-data.frame(actual = valid.df.norm$Car_Cancellation, predicted = logit.reg.pred.valid, errors =  logit.reg.pred.valid - valid.df.norm$Car_Cancellation)


#Produce a confusion matrix
confusionMatrix(as.factor(ifelse(logit.reg.pred.train > 0.5, 1, 0)), as.factor(train.df.norm$Car_Cancellation))

confusionMatrix(as.factor(ifelse(logit.reg.pred.valid> 0.5, 1, 0)), as.factor(valid.df.norm$Car_Cancellation))

```

```{r}
#
gain.reg.train <-gains(train.df.norm$Car_Cancellation, logit.reg.pred.train, groups=10)
print.gains(gain.reg.train)

gain.reg.vaild <-gains(valid.df.norm$Car_Cancellation, logit.reg.pred.valid, groups=10)
print.gains(gain.reg.vaild)

#Plot lift chart for train
plot(c(0,gain.reg.train$cume.pct.of.total*sum(train.df.norm$Car_Cancellation))~c(0,gain.reg.train$cume.obs),xlab="# cases", ylab="Cumulative", main="Plot lift chart for logistic regression(train.df.norm)", type="l")
lines(c(0,sum(train.df.norm$Car_Cancellation))~c(0, dim(train.df.norm)[1]), lty=2)

#Plot a Decile chart for train
heights <-gain.reg.train$mean.resp/mean(train.df.norm$Car_Cancellation)
midpoints <- barplot(heights, names.arg = gain.reg.train$depth, ylim = c(0,9),
xlab = "Percentile", ylab = "Mean Response", main = "Decile wise lift chart for logistic regression(train.df.norm)")
text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.8)


#Plot lift chart for valid
plot(c(0,gain.reg.vaild $cume.pct.of.total*sum(valid.df.norm$Car_Cancellation))~c(0,gain.reg.vaild $cume.obs),
     xlab="# cases", ylab="Cumulative", main="Plot lift chart for logistic regression(valid.df.norm)", type="l")
lines(c(0,sum(valid.df.norm$Car_Cancellation))~c(0, dim(valid.df.norm)[1]), lty=2)

#Plot a Decile chart for valid
heights <-gain.reg.vaild$mean.resp/mean(valid.df.norm$Car_Cancellation)
midpoints <- barplot(heights, names.arg = val.reg.valid$depth, ylim = c(0,9),
xlab = "Percentile", ylab = "Mean Response", main = "Decile wise lift chart for logistic regression(valid.df.norm)")
text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.8)

```


# Model_2 - Classification Tree
```{r}
#Grow a default tree
Cancellation.ct <- rpart(Car_Cancellation ~. , data = train.s.df, method = "class")
Cancellation.ct

prp(Cancellation.ct, type = 1, extra = 1, split.font = 1, varlen = -10)

```

```{r}
#Grow an “ideal” size tree

#1. Create a deep tree by setting parameter cp=0.00001, to minimize pruning, and add parameter xval=5 to use cross-validation to estimate the tree error rate.
Cancellation.ct.2 <- rpart(as.factor(Car_Cancellation) ~., data = train.s.df,
                    method = "class", cp = 0.00001, minsplit = 5, xval = 5)
#2. Print tree attribute cptable, to show a table with results recorded during the tree-growing process.
printcp(Cancellation.ct.2)
plotcp(Cancellation.ct.2)

#when cp=0.00403769, the minimize xerror 
#3. Prune the tree to the best level by setting CP to the value that minimizes xerror, namely, cp=0.00403769 

pruned.ct <- prune(Cancellation.ct.2 , cp = 0.00403769 )
pruned.ct

prp(pruned.ct, type = 1, extra = 1, split.font = 1, varlen = -10)

```

## Model_2 -  Model validation 
```{r}
#Predict the CLASS for a specific data set
#for train.s.df
pred.train <-predict(pruned.ct,train.s.df,type = "class")
confusionMatrix(pred.train,factor(train.s.df$Car_Cancellation))

#for valid.df
pred.valid <- predict(pruned.ct,valid.df,type = "class") 
confusionMatrix(pred.valid, as.factor(valid.df$Car_Cancellation))

# add a new column
train.s.df$cancel <- as.numeric(pred.train)
valid.df$cancel <- as.numeric(pred.valid)
```

```{r}
#
gain.train <- gains(train.s.df$Car_Cancellation,train.s.df$cancel, groups=10)
print.gains(gain.train )

gain.valid<-gains(valid.df$Car_Cancellation,valid.df$cancel, groups=10)
print.gains(gain.valid)

#Plot a Gains chart for train.s.df
plot(c(0,gain.train$cume.pct.of.total*sum(train.s.df$Car_Cancellation))~c(0,gain.train$cume.obs), xlab="# cases", ylab="Cumulative", main="Gains chart for classification tree(train.s.df)", type="l")
lines(c(0,sum(train.s.df$Car_Cancellation))~c(0, dim(train.s.df)[1]), lty=2)

#Plot a Decile chart for train.s.df
heights <- gain.train$mean.resp/mean(train.s.df$Car_Cancellation)
midpoints <- barplot(heights, names.arg =gain.train$depth, ylim = c(0,12),
xlab = "Percentile", ylab = "Mean Response", main = "Decile-wise lift chart for classification tree(train.s.df)") 
text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.9)

#Plot a Gains chart for valid.df
plot(c(0,gain.valid$cume.pct.of.total*sum(valid.df$Car_Cancellation))~c(0,gain.valid$cume.obs), xlab="# cases", ylab="Cumulative", main="Gains chart for classification tree(valid.df)", type="l")
lines(c(0,sum(valid.df$Car_Cancellation))~c(0, dim(valid.df)[1]), lty=2)

#Plot a Decile chart for valid.df
heights <- gain.valid$mean.resp/mean(valid.df$Car_Cancellation)
midpoints <- barplot(heights, names.arg = gain.valid$depth, ylim = c(0,12),
xlab = "Percentile", ylab = "Mean Response", main = "Decile-wise lift chart for classification tree(valid.df)") 
text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.9)

```


# Model_3 - Random Forest

```{r}
# Modal_2- Random Forest
set.seed(123)

rf <- randomForest(as.factor(Car_Cancellation) ~  
                     Longdistance + 
                     PointToPoint + 
                     HourlyRental + 
                     month+
                     wkday+
                     Waiting + 
                     online_booking + 
                     mobile_site_booking + 
                     traditional_booking + 
                     Dist_trip_length, data = train.df, proximity = TRUE)


tree_pre <- predict(rf, valid.df, type = "response")

print(rf)

print(importance(rf))

### data distribution for user_id
as.data.frame(table((df_1$user_id),df_1$Car_Cancellation)) -> user_data

user_data <- as.data.frame.matrix(table(df_1$user_id,df_1$Car_Cancellation))
user_data$user.id <-rownames(user_data)
user_table <- data.frame(user_data) 

###Variable importance plot
varImpPlot(rf,
           sort = T,
           n.var =6,  
           main = "Variable Importance")
?varImpPlot

```

## Model_3 - Model validation  
```{r}
#confusionMatrix
confusionMatrix(as.factor(tree_pre),as.factor(valid.df$Car_Cancellation))

#Gain chart
pred.prob.valid <- predict(rf,valid.df,type = "prob")
 
gain.valid <- gains(valid.df$Car_Cancellation, pred.prob.valid[,2], groups=10)
print.gains(gain.valid) 

## Validation Lift Charts
plot(c(0,gain.valid$cume.pct.of.total*sum(valid.df$Car_Cancellation))~c(0,gain.valid$cume.obs),
     xlab="# cases", ylab="Cumulative", main="Gains chart for random forest(valid.df)", type="l")
lines(c(0,sum(valid.df$Car_Cancellation))~c(0, dim(valid.df)[1]), lty=2)

#Decile-wise lift chart
heights <- gain.valid$mean.resp/mean(valid.df$Car_Cancellation)
midpoints <- barplot(heights, names.arg = gain.valid$depth, ylim = c(0,9),
 xlab = "Percentile", ylab = "Mean Response",
main = "Decile-wise lift chart for random forest(valid.df)")

text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.8)


```





